---
title: "R Notebook"
output: html_notebook
---

Content for the scrooge poster

```{r}
set.seed(42)

library(tidyverse)
library(broom)
library(scales)
library(GUM)
library(forcats)
library(stringr)
library(ggrepel)
library(ggthemes)
library(lazyeval)
library(trelliscopejs)
library(LBSPR)
library(simfish)
library(ggsci)
library(hrbrthemes)
library(viridis)
library(ggExtra)
library(ggalt)
library(LIME)

hrbrthemes::update_geom_font_defaults()

poster_theme <- theme_ipsum(base_size = 32, plot_title_size = 38,axis_title_size = 38,
                            strip_text_size = 35, strip_text_face = 'bold')

theme_set(poster_theme)

dat <- data.table::fread('~/Box Sync/Shared Upside Results/ProjectionData.csv',
                         showProgress = T)  %>% 
  set_names(colnames(.) %>% tolower())


a = dat %>% 
  group_by(policy,year, scenario) %>% 
  summarise(tc = sum(catch, na.rm = T), tmsy = sum(msy)) %>% 
  filter(year == 2010 | year == 2050) %>% 
  ungroup() %>% 
  mutate(pc = tc / tc[policy == 'Historic'] - 1)


a <- dat %>% 
  filter(year == 2050, policy == 'Fmsy', 
         scenario == 'All Stocks') %>% 
  mutate(c_v_msy = catch / msy)

a = dat %>% 
  filter(idorig == 'WGSSDS-SOLECS-1970-2011-NEUBAUER', policy == 'Fmsy', 
         scenario == 'All Stocks') %>% 
  ggplot(aes(year,bvbmsy)) + 
  geom_line()
  


dat %>% 
  group_by(policy,year, scenario) %>% 
  summarise(p_overfished = mean(bvbmsy < 1),
            p_overfishing = mean(fvfmsy > 1)) %>% 
  filter(year == 2010 | year == 2050) 
  

  


ram <- dat %>% 
  filter(dbase =='RAM')

current_ram <- ram %>% 
  group_by(year) %>% 
  summarise(tc = sum(catch, na.rm = T)) %>% 
  filter(year == 2012) %>% {
    .$tc
  }

global_catch <- 81e6

percent_unassessed <- ((1 - current_ram / global_catch) * 100) %>% round(0)


```


# Figure 1

```{r}
strain_lbspr <- function(sigma_recruits = 0, sigma_f = 0, ac_recruits = 0,HistoricalF = 0.1){

  
    vbk = 0.21
    linf = 64.58
    lwa = 0.0245
    lwb = 2.79
    S50 = 30
    selex_input = "length"
    M50 = 34
    maturity_input = "length"
    binwidth = 1
    CVlen = 0.1
    SigmaR = sigma_recruits
    M = 0.2
    rho = ac_recruits
  
  
  lh <-
  create_lh_list(
    vbk = vbk,
    linf = linf,
    lwa = lwa,
    lwb = lwb,
    S50 = S50,
    selex_input = selex_input,
    M50 = M50,
    maturity_input = maturity_input,
    binwidth = binwidth,
    CVlen = CVlen,
    SigmaR = SigmaR,
    M = M,
    rho = rho,
    F1 = HistoricalF,
    SigmaF = sigma_f
  )
  
  simmed_fish <- sim_pop(
  lh = lh,
  Nyears = 50,
  Fdynamics = 'Constant',
  Rdynamics = 'BH',
  nburn = 5,
  modname = 'blah',
  seed = 42,
  init_depl = 0,
  comp_sample = 1000,
  pool = T,
  Nyears_comp = 50,
  mismatch = F
)
  
  lbspr_lengths <- simmed_fish$LF %>%
  t() %>% 
  as_data_frame() %>% 
  set_names(1:dim(.)[2])

lbspr_lengths <- data.frame(LMids = simmed_fish$mids %>% as.numeric(),lbspr_lengths, check.names = F)


babel_lengths <- new("LB_lengths")

babel_lengths@LMids <- simmed_fish$mids

babel_lengths@LData <- lbspr_lengths %>% as.matrix()
  
  babel_lengths@Years <-
  as.numeric(lbspr_lengths %>% select(-LMids) %>% colnames())

babel_lengths@NYears <- length(babel_lengths@Years)

babel_pars <- new("LB_pars")

babel_pars@Species <- "babelfish"

babel_pars@Linf <- linf

babel_pars@L50 <- M50

babel_pars@L95 <-  M50+1

babel_pars@MK <- M / vbk

babel_pars@BinWidth <- 1

babel_fit <- LBSPRfit(babel_pars, babel_lengths,
                      yrs = 1:babel_lengths@NYears, Control = list(modtype = 'absel',
                                                                   Nage = lh$AgeMax
                                                                   ))
real <- data_frame(year = 1:50, spr = simmed_fish$SPR_t, name = 'True Value')

lbspr <- data_frame(year = 1:50, spr = babel_fit@SPR, name = 'LBSPR Estimate')

spr_comp_plot <- real %>%
  bind_rows(lbspr) %>% 
  ggplot(aes(year,spr, color = name)) +
  geom_line(size = 2) +
  scale_y_continuous(limits = c(0,NA))

real <- data_frame(year = 1:50, fvm = simmed_fish$F_t / lh$M, name = 'True Value')

lbspr <- data_frame(year = 1:50, fvm = babel_fit@FM, name = 'LBSPR Estimate')

fvm_comp_plot <- real %>%
  bind_rows(lbspr) %>% 
  ggplot(aes(year,fvm, color = name)) +
  geom_line(size = 2) +
  scale_y_continuous(limits = c(0,NA))

return(list(spr_comp_plot = spr_comp_plot,fvm_comp_plot = fvm_comp_plot))
}

f <- 0.3

no_strain <- strain_lbspr(HistoricalF = f)

no_strain$spr_comp_plot

no_strain$fvm_comp_plot

plenty_strain <- strain_lbspr(sigma_recruits = 1.1, ac_recruits = 0.75,
                              HistoricalF = f, sigma_f = 0)

plenty_strain$spr_comp_plot

spr_strain_plot <- plenty_strain$spr_comp_plot  + 
  xlab('Time') + 
  ylab('Spawning Potential Ratio') + 
  scale_color_npg(guide = F) + 
  labs(title = 'A: Rec. Assumptions Violated')

spr_no_strain_plot <- no_strain$spr_comp_plot  + 
  xlab('') + 
  theme(axis.title.y = element_blank()) + 
  labs(title = 'B: Rec. Assumptions Satisfied') + 
  scale_color_npg(name = '', labels = c('LBSPR Estimate','True Value'))

lbspr_performance_plot <- gridExtra::arrangeGrob(spr_strain_plot,spr_no_strain_plot,
                                                  ncol = 2, nrow = 1)
png(file = 'test.png', width = 20, height = 12, units = 'in', res = 72)
 gridExtra::grid.arrange(spr_strain_plot,spr_no_strain_plot,
                                                  ncol = 2, nrow = 1,
                         widths = unit(c(10,10),'in'),heights = unit(c(12),'in'))
 dev.off()


# 
# cowplot::plot_grid(spr_strain_plot,spr_no_strain_plot, labels = 'AUTO')
# 
# cowplot::save_plot('test.png', lbspr_performance_plot, base_height = 12,
#                    base_aspect_ratio = 1.25)

```


```{r}

r <- 1.1

k <- 100

q <- .2

alpha <- 1e-2

p = 5
 
cost = 20

beta = 1.3

msy <- r*k / 4

bmsy <- k / 2

emsy <- msy / (q * bmsy)

pmsy <- p * msy - cost * emsy ^ beta

oa <- data_frame(year = 1965:1990, b = NA, f = NA, catch = NA, profits = NA, effort = NA)

time <- length(1965:1990)

oa$b[1] <- k

oa$effort[1] <- .01

oa$f[1] <- q * oa$effort[1]

oa$catch[1] <- oa$f[1] *oa$b[1] 

oa$profits[1] <- (p * oa$catch[1]  - cost * oa$effort[1] ^ beta) / pmsy

for (i in 2:time) {
  
  last_b <- oa$b[i - 1]
  
  oa$b[i] <- last_b + last_b * r * (1 - last_b / k) - oa$catch[i - 1]
  
  oa$effort[i] <- oa$effort[ i - 1] + alpha * oa$profits[ i - 1]
  
  oa$f[i] <- oa$effort[i] * q
  
  oa$catch[i] <- oa$b[i] * oa$f[i]
  
  oa$profits[i] <-  p * oa$catch[i] - cost * oa$effort[i] ^ beta
  
}

# oa  %>% 
#   mutate(commname = 'Tragic Hardinfish', bvbmsy = b / bmsy, 
#          fvfmsy = f / (r/2)) %>% 
#   select(commname, year, bvbmsy, fvfmsy) %>% 
#   ggplot(aes(bvbmsy, fvfmsy)) + 
#   geom_hline(aes(yintercept = 1), linetype = 2) + 
#   geom_vline(aes(xintercept = 1), linetype = 2) +
#   geom_path() + 
#   geom_point(aes(fill = year), shape = 21, size = 10) + 
#   scale_fill_viridis(name = 'Year') + 
#   labs(x = 'B/Bmsy', y = 'F/Fmsy') + 
#   facet_wrap(~commname) +
#   theme(legend.key.height = unit(1.5,'in'),
#         legend.key.width = unit(.35,'in'))

# Ram example

historic_ram <- ram %>% 
  filter(idorig == 'NAFO-SC-YELL3LNO-1960-2009-BAUM', year <= 1990) %>% 
  select(commname, year, bvbmsy, fvfmsy) %>% 
  mutate(commname = paste('B) RAM Data - ',commname, sep = ''))


oa_plot <- oa  %>% 
  mutate(commname = 'A) Stylized Model - Tragic Hardinfish', bvbmsy = b / bmsy, 
         fvfmsy = f / (r/2)) %>% 
  select(commname, year, bvbmsy, fvfmsy) %>% 
  bind_rows(historic_ram) %>% 
  ggplot(aes(bvbmsy, fvfmsy)) + 
  geom_hline(aes(yintercept = 1), linetype = 2) + 
  geom_vline(aes(xintercept = 1), linetype = 2) +
  geom_path() + 
  geom_point(aes(fill = year), shape = 21, size = 10) + 
  scale_fill_viridis(name = 'Year') + 
  labs(x = 'B/Bmsy', y = 'F/Fmsy') + 
  facet_wrap(~commname) +
  theme(legend.key.height = unit(1.5,'in'),
        legend.key.width = unit(.35,'in'))

oa_plot

ggplot2::ggsave('oa_plot.png', oa_plot, width = 20,height = 12, unit = 'in')

```



# Project Summary

Can Economic Data Improve Data-Limited Stock Assessments?

# The Problem






# Project Summary

Most fisheries around the world require management guidance but lack the robust data streams required for state-of-the-art fisheries management. To resolve this problem, a large and growing suite of "data-limited stock assessments" (DLAs) have emerged, designed to provide management advice using relatively minimal data (but generous assumptions). Interestingly though, nearly all of the quantitative DLAs established rely exclusively on fish-centric data, for example length frequency distributions [e.g. @Gedamke2006; @Hoenig1983; @Hordyk2016], catch data [e.g. models summarized in @Rosenberg2017; @Carruthers2014]. While these data are clearly critical to proper understanding of a fishery's status, they also ignore the potential light that the economic history of a fishery may shed on its current biological status. This project proposes to address this challenge, by developing a quantitative method for integrating economic information into the stock assessment process, which we can loosely term economic-DLAs (eDLAs).


DLAs common

Big problems when assumptions are violated

Graph showing LBSPR thing

Graph of problem with data from central coast

# Economics and Fisheries

Summaries of fleet dynamics and theory behind linkages of profits and fishing effort

Maybe some plots showing how different incentives create different behavior

# Proposed Methods

Empirical Examination

Very few papers actually making this link 

How do we propose to fit this? Test out of sample predictive ability of different fleet model structures and parameters 

Simulation testing!

Application to California!

# Expected outcomes

New source of data for DLAs

Isolating recruitment from F

# Results

Example figure of open access looking thing from RAM

```{r}


ram <- dat %>% filter(dbase == 'RAM' & year <= 1990, is.na(fvfmsy) == F)

ram_series <- ram %>%
  select(idorig, year, bvbmsy, fvfmsy, catch) %>%
  nest(-idorig) %>%
  mutate(oa = map_plot(
  data,
  ~ ggplot(.x, aes(bvbmsy, fvfmsy)) + geom_vline(aes(xintercept = 1), linetype = 2) + geom_hline(aes(yintercept = 1), linetype = 2) + geom_path()  + geom_point(aes(color = year))
  ))
# 
# trelliscope(ram_series %>% select(idorig,oa), name = 'blah', self_contained = T,panel_col = 'oa')



```

price correlations
```{r}

edat <- read_csv(file.path('../data',
                           'SouthCoastHumanActivitiesCommercialFisheriesEconomicandSpatialData1992to2012',
                           'sc_comm_fisheries_region_landings_data_1992_2012.csv')) %>% 
  set_names(colnames(.) %>% tolower()) %>% 
  mutate(pounds = str_replace(pounds,',','') %>% as.numeric(),
         revenue = str_replace_all(revenue,'[\\$|,]','') %>% as.numeric(),
         `average price` = str_replace(`average price`, '[\\$]','') %>% as.numeric(),
         `count of landings` = as.numeric(`count of landings`)
         )


x <- function(a,f,d) {
  
  f(a,d)
  
}

x(c(1,2,3),lag,1)

leads_and_lags <- function(df,variable,foo, x) {
  foo_name <- enquo(foo)
  
  for (i in seq_along(x)) {
  name <- paste(quo_name(foo_name),x[i], sep = '_')
  
 df <-  mutate(df, 
         !!name := foo(!!variable,x[i])
         )
  }
  
  return(df)
}

edat <- edat %>% 
  group_by(region,fishery,gear) %>% 
  arrange(year) %>% 
  leads_and_lags(quo(`average price`),lag,0:5) %>% 
  leads_and_lags(quo(`average price`),lead,1:5)


price_pounds <- edat %>% 
  select(region,fishery,gear,year,pounds,contains('lead'), contains('lag')) %>% 
  gather('variable','value',contains('_')) %>% 
  group_by(region,fishery,gear,variable) %>% 
  summarise(pearson_correlation = cor(pounds,value, use = 'na.or.complete'),
            cross_correlation = ccf(pounds,value,lag.max = 0, plot = F, na.action = na.pass)$acf %>% as.numeric()) %>% 
  mutate(is_lag = str_detect(variable,'lag'),
         lead = str_split(variable,'_', simplify = T)[,2] %>% as.numeric()) %>% 
  mutate(lead = ifelse(is_lag == T, -lead,lead))

price_pounds <- price_pounds %>% 
  ungroup() %>% 
  mutate(fishery = tools::toTitleCase(fishery %>% tolower()),
         gear =tools::toTitleCase(gear %>% tolower())) %>% 
  mutate(fish_gear = paste(fishery,gear, sep = '-'))


price_pounds$fishery[price_pounds$fishery == 'Nearshore'] <- 'Nearshore Finfish'



good_examples <- c('California Halibut-Hook & Line',
                   'Nearshore Finfish-Trap',
                   'Spot Prawn-Trap',
                   'Lobster-Trap')

price_pound_corr_plot <- price_pounds %>% 
  filter(is_lag == T,
         is.na(region) == F, is.na(fishery) == F, is.na(gear) == F) %>%
  filter(fish_gear %in% good_examples) %>% 
  ggplot(aes(lead,cross_correlation)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_lollipop(aes(color = fishery), size = 4) + 
  scale_color_npg(guide = F) +
  facet_wrap( ~ fish_gear) +
 labs(x= 'Price Lag', y = 'Price-Catch Cross-Correlation')
  
ggplot2::ggsave('price-correlations.png', price_pound_corr_plot, width = 25,height = 12, unit = 'in')







```

Maybe 

IMARPE-PANCHPERUNC-1959-2008-HIVELY
AFSC-POPERCHGA-1959-2010-STACHURA

AFWG-GOLDREDNEAR-1983-2010-NEUBAUER

IMARPE-PANCHPERUNC-1959-2008-HIVELY

NAFO-SC-YELL3LNO-1960-2009-BAUM * good one

NEFSC-HADGB-1931-2010-CHING

NEFSC-HERRNWATLC-1960-2005-OVERHOLTZ

NZMFishLOBSTERWG-RROCKLOBSTERCRA1-1945-2001-JENSEN

NAFO-SC-YELL3LNO-1960-2009-BAUM * very good one

NEFSC-HADGB-1931-2010-CHING

NWWG-CODICE-1952-2012-CHING

NZMFishLOBSTERWG-RROCKLOBSTERCRA1-1945-2001-JENSEN



ICCAT-BMARLINATL-1956-2010-CHING




